Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator

### 2023-09-11 ms.qmd
Figure and heterogeneity

The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units.

@@

To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`). The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]`, a not ignorable difference.

>>>>

The model results confirmed the visual impressions described above: @tbl-avg indicates, with great certainty, that mood increased by about [`r paste0(map(h_avg_$hypothesis[3,4:5], ~number(.x, .001)), collapse = ", ")`] units (@fig-coef, "Raw").

@@

To further interpret the magnitude of the difference in moods between pre- and during play, we compared the contrast to the natural variations in mood as estimated by the model (standard deviation of the between-person intercept heterogeneity ($\tau_{\alpha_0}$ = `r posterior_pop_sum$x[1]`), between-person pre-during difference heterogeneity ($\tau_{\alpha_1}$ = `r posterior_pop_sum$x[2]`) and to the residual standard deviation ($\sigma$ = `r posterior_pop_sum$x[3]`). The standardized (with the total residual variation from the three above components) pre-play -- during play contrast was `r posterior_pop_sum$x[6]` (@fig-coef, "Scaled"), a not ignorable difference.

```{r}
#| label: heterogeneity-calculate

posterior_pid <- coef(fit, summary = FALSE)$pid[,,"post1"]
posterior_pid_scaled <- apply(
  posterior_pid, 2, 
  function(x) x / posterior_pop$total_residual
) |> 
  apply(2, function(x) mean_qi(x)) |> 
  bind_rows(.id = "pid")
posterior_pid <- posterior_pid |> 
  apply(2, function(x) mean_qi(x)) |> 
  bind_rows(.id = "pid")

posterior_pid_sum <- bind_rows(
  "b_post1" = posterior_pid,
  "b_post1_scaled" = posterior_pid_scaled,
  .id = "name"
)

x <- posterior_pop |> 
  mutate(
    prop_pos = pnorm(0, b_post1, sd_pid__post1, lower.tail = FALSE)
  ) |> 
  mean_qi(prop_pos) |> 
  mutate(x = str_glue("{percent(prop_pos, .1)} [{number(.lower*100, .1)}, {number(.upper*100, .1)}]")) |> 
  pull(x)
```

We then turned our attention to between-person differences in the difference in during and pre-play mood. With a sample of `r s$n_sub`, it is not surprising that the population-level contrast is credibly different from zero. We therefore asked the subsequent question: What proportion of the population from which the sample was drawn would we expect there to be an elevation in mood during play? The boxplots in @fig-coef show the distribution of player-specific contrasts estimated by the model, which indicated that `r x` of players in the population experience a positive shift in mood during play, compared to their moods during login.

```{r}
#| label: fig-coef
#| include: true
#| fig.width: 3
#| fig.height: 1.4
#| fig.cap: "Difference in mood during play compared to mood at session start. Points and intervals indicate posterior means and 95%CIs of the population level difference; boxplots summarize the player-level estimates. Raw estimates are on the VAS scale, whereas Scaled estimates indicate contrasts divided by the total residual deviance of the model."

posterior_pop |> 
  select(starts_with("b_")) |> 
  pivot_longer(everything()) |>
  group_by(name) |> 
  mean_qi(value, .width = 0.95) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "During play - pre-play contrast",
    breaks = extended_breaks(7)
  ) +
  scale_y_discrete(
    breaks = c("b_post1", "b_post1_scaled"),
    labels = c("Raw", "Scaled")
  ) +
  coord_cartesian(xlim = c(-0.25, 0.65)) +
  geom_vline(xintercept = 0, linewidth = .33, col = "dodgerblue") +
  geom_boxploth(
    data = posterior_pid_sum,
    aes(x = y, group = name), 
    outlier.shape = NA,
    notch = TRUE,
    size = 0.25,
    width = .2
  ) +
  geom_pointrangeh(
    aes(xmin = .lower, xmax = .upper),
    fatten = 2.5
  ) +
  theme(
    axis.title.y = element_blank()
  )
```
